Debugging by {\huge\bf\textit{lastChange}}

### 2011-03-11 papers/fse2011/sig-alternate.tex

\subsection{{\large\bf\textit{lastChange}} on object property}
To simplify the algorithm explanation and defer technical details, we
define two basic operations and later we explain the details of these
two operations. The first operation is \texttt{objectId()}: given a
JavaScript object it returns an integer as its identifier. This
identifier is unique to the object during one execution.  By using an
object id instead of an object reference we allow the garbage
collector to reclaim the space for dead objects just as it would in the
absence of the debugger. The second
operation is \texttt{setPropertyChangeHook()}: given a function and a
string, the function is called whenever a property changes and its
name matches the string. For example, if the string is 
\texttt{foo}, changes to \texttt{bar.foo} or \texttt{baz.foo} would
call the function.  The callback function receives a reference to the
owner of \texttt{foo}.

@@

\subsection{{\large\bf\textit{lastChange}} on variable} 
In JavaScript, every frame has a scope chain and every available
variable in the frame comes from one of the scopes in the frame's
scope chain. Once the developer asks for the last change of a variable with name 
\texttt{foo} at the reproduction point, the debugger first determines the
variable's scope as follows: it iterates over the scopes in the scope
chain and the first scope which has a variable with the same name is
the variable's scope. There are five different scope types: global, local,
closure, \texttt{with} and \texttt{catch}. We explain these
cases in two groups.

@@

\begin{figure}[htp]
\centering 
%\begin{algorithm}
\begin{frame}
\begin{algorithmic}

\FOR{$q$ in $lastChange$ queries} 
 \FOR {$p$ is defined at the $q$ result}
   \IF {$p$ is a $lastChange$ on object property} 
     \STATE the property {\bf owner id} must be stored at $q$ change events. 
   \ELSIF {$p$ is a $lastChange$ on variable} 
     \STATE the variable {\bf scope id} must be stored at $q$ change events.
	 \ENDIF 
 \ENDFOR 
\ENDFOR

\end{algorithmic}
\end{frame}
%\end{algorithm}
\caption{\textit{lastChange} queries dependency analysis.}
\label{fig:dependency-analysis}
\end{figure}
%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented\footnote[1]{http://code.google.com/p/querypoint-debugging} it in an extension to the Firebug
JavaScript debugger\footnote[2]{http://getfirebug.com}. %\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser. %\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface \cite{JSD} and 
\textit{Querypoint} is developed over this interface. Our prototype implements the four primitive operations in JavaScript 
using techniques which are cumbersome and comparatively slow to execute. However the JavaScript prototype is
easy to explore, change, and share with others for feedback.
  A professionally useful debugger would implement these primitive operations within the JavaScript engine.

\subsection{{\normalsize\bf\texttt{objectId()}} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. 
%\footnote[3]{This use of a regular object property  might change
%the program behaviour. For example it will appear as an extra property in 
%\texttt{for(property in object)} loops. 
%The next version of Firefox (https://developer.mozilla.org/en/JavaScript/Reference\\/Global\_Objects/Object/defineProperty) provides \texttt{defineProperty}, 
%allowing a property to be added which is invisible to enumeration.} Note that \texttt{\_objectId} is
%not set for all objects but only those objects that need an id.

\subsection{{\normalsize\bf\texttt{setPropertyChangeHook()}} operation}
The Firefox JavaScript engine supports watching property changes in an
object. Every object has a function \texttt{watch(propName, callback)}
which receives two parameters, a property name and function.  Whenever
the property with the given name changes, the \texttt{callback} function is
called. The hook set by this function remains enabled even if the
property is deleted and defined again.

@@

The only data we have is the object creation location including the
file url and the line number and the goal is to get a reference to
this object. Although in most regular cases we have only one
statement and one object creation in a line, there are cases that more
than one object might be created in a line. There is no simple way to
recognize the interesting object among these new objects. So instead
of one object, we monitor all new objects created in the line.

An object might be created by one of these statements: object
initializer (\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statment (\texttt{function()}). 
By parsing the source code we can recognize the statements that
create an object. The next step is getting a reference to the new object.

@@

The new object can also be set as the property of a parent object by a
colon (\texttt{:}). This case is also treated similar to previous case. The
only difference is that the full path of property from the root parent
in the local scope should be considered as the assignee statement. For
example in Figure~\ref{fig:objectCreation}(b), if the creation line
is 20, we keep \texttt{parentObject.newObject}. In this case stepping
should be continued until the end of line 22 for getting a reference
to the created new object.

There is another case for new objects where they are passed as
arguments to other functions (Figure~\ref{fig:objectCreation}(c)). In these
cases instead of step-over we use step-in and we get the corresponding
argument as a new object.

Although this approach is successful in many ordinary cases, there are some 
cases that a more comprehensive analysis needed for correct behaviour. For example in a case
where the new object is assigned to \texttt{a[++i]} where i is a variable
(Figure~\ref{fig:objectCreation}(d)), \texttt{a[++i]} is kept in the list
of assignee statements. Obviously, evaluating this statement doesn't return
a reference to the new object. Our prototype implementation does not handle these kinds of unusual cases yet.

Throughout this section we implicitly assumed that the object will be created at 
the same location in the next execution. So, what happens if it
is not true? Once the execution reaches the reproduction point, it
reveals that the object has been created in different location. This time,
prototype re-executes the program considering both locations as possible 
object creation locations.

@@

(a) Case 1:
 20 x.y = new myConstructor();

(b) Case 2:
 19  parentObject = {
 20   newObject : {
 21    x : 5
 22  }}

@@

(d) Case 4:
 20 a[++i] = new myConstructor();
\end{lstlisting} 
%\end{verbatim}
\caption{Examples of different cases in getting references to created JavaScript objects.}
\label{fig:objectCreation}
\end{figure}

\subsection{{\normalsize\bf\texttt{scopeId()}} operation}
The prototype sets a breakpoint at the beginning of all code blocks needing
a scope id. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook is hit,  meaning a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval("var \_scopeId = 10")} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{{\normalsize\bf\texttt{setVariableChangeHook()}} operation}
Variables defined in local, closure, and catch scopes are only changed in
their scope; that includes the defining scope and scopes nested whithin them.
For example in Figure~\ref{fig:js-closure}, variable \texttt{myVar} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line where the variable is defined. In Figure
~\ref{fig:js-closure}, if the execution is paused at line 17 and the
last change of \texttt{myVar} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different: \texttt{myVar} in 
\texttt{firstChild} is a local variable
but in \texttt{secondChild} it is a closure variable.

@@

10  function parent(){
11    var myVar;
12    myVar = ...;
13    function myfun(){
14      myVar = ...;
15      function firstChild(){
16        var myVar;
17        myVar = ...;
18      }  
19      function secondChild(){
20        myVar = ...;			      
21      }
22    }  
23  }    
\end{lstlisting}
%\end{verbatim}
\caption{Sample JavaScript code demonstrating local and clousure variables.}
\label{fig:js-closure}
\end{figure}

\subsection{Re-Execution, reproduction point and data collection}
\textit{Querypoint} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypoint}
is also able to automatically create them from the first execution.

@@

Since we don't need to compare values across executions to implement \textit{lastChange}, we can get more information if we do compare:
 different points of last change on different executions will signal that the execution is not deterministic. (Note the converse is not true, many non-deterministic 
programs will give identical results for \textit{lastChange}.)
Consider the example in Figure~\ref{fig:nondeterministic-values}. 
Assume that at the first execution, developers sees that \texttt{a} is \texttt{true} and asks
for the \textit{lastChange} of \texttt{a} at the breakpoint. In the re-exeuction, \texttt{a} is \texttt{false},
and \texttt{b} is \texttt{true} at the reproduction point. It means that \texttt{lastChange}
result shows line 10, which is a correct answer for this execution but not for the previous one.
When user asks for the \textit{lastChange} on variable \texttt{a}, debugger stores \texttt{a} value and compares
it to the \texttt{a} value in the next executions. So if this value is different, debugger informs the developer
that the \textit{lastChange} query is made on a different value.
% In cases that the value is not a primitive but an object,
%the primitive properties of the object examined by the developer are stored and compared.

@@

10 a = false;
11 b = false;
...
20 if ( random() ) { a = true; } else { b = true; }
...
30 if (a || b) { bug();  /* breakpoint */ }
\end{lstlisting}
%\end{verbatim}
\caption{\textit{lastChange} on non-deterministic values.}
\label{fig:nondeterministic-values}
\end{figure}

\subsection{Combination of {\large\bf\textit{lastChange}} and breakpoint debugging}
\label{sec:pausing}
Using \textit{lastChange} a developer can work backwards on the flow of data, but sometimes bugs are more obvious when
we watch control flow forwards. Consider for example,

@@

In the case that the execution is deterministic, the \textit{lastChange}
event index can be used as an index for a conditional breakpoint \cite{Boothe, Maruyama}. 
Moreover, this works in a non-deterministic execution if non-determinism
has no effect on the event index.  The debugger can easily set this conditional breakpoint
and reexecute the program. If the stream of change events differs in this re-execution the
developer can be warned that the conditional breakpoint may not be the \textit{lastChange} point.
However, if the stream of events is the same, we do not know that the 
conditional breakpoint matches the last 
change, this fact  can only verify  at the reproduction point. These theoretical concerns are not
likely to cause significant practical problems: if the value at the conditional breakpoint is not suspcious, 
then the developer will know to simply return to \textit{lastChange} or move on to another tactic to find the bug.

@@

We designed the example programs to require two steps of \textit{lastChange}; 
all four users took more steps (9, 13, 12, and more than 30) to find the defect location with breakpoints.  
To find locations for breakpoints two users scrolled through the source, 
another searched the text, a third set a lot of breakpoints to get an understand of the control flow. 
Based on our own experience we expect these strategies represent the kinds of approaches 
developers have available. These operations are time consuming and tedious. In contrast, all four users 
found the defect location with just two \textit{lastChange} operations.
% Based on our measurements,
%this makes \textit{lastChange} about three times faster than breakpoints for the navigation from 
%the reproduction point to the defect
We recognize that these programs were designed to highlight lastChange and many kinds of debugging 
issues have been hidden by the design of our tests. Nevertheless our results show that, when a defect 
relates to incorrect values and a developer recognizes this, then the operational 
mechanics of lastChange lead to the defect much more quickly than breakpoints.

@@

@@

%In our own experience, \textit{lastChange} dramatically reduces the work in
%setting and removing breakpoints, so work to cause re-execution  become noticable.
%Therefore automatic mechanisms
%for re-execution will be corresponding more valuable for debugging
%with \textit{lastChange} (Our \textit{Querypoint} prototype  
%implements a simple form of automatic 
%re-execution for both breakpoint and \textit{lastChange}).

>>>>

\subsection{{\large\bf\textit{lastChange}} on Object Property}
To simplify the algorithm explanation and defer technical details, we
define two basic operations and later we explain the details of these
two operations. The first operation is \texttt{objectId()}: given a
JavaScript object it returns an integer as its identifier. This
identifier is unique to the object during one execution.  By using an
object id instead of an object reference we allow the garbage
collector to reclaim the space for dead objects just as it would in the
absence of the debugger. The second
operation is \texttt{setPropertyChangeHook()}: given a function and a
string, the function is called whenever a property changes and its
name matches the string. For example, if the string is 
\texttt{foo}, changes to \texttt{bar.foo} or \texttt{baz.foo} would
call the function.  The callback function receives a reference to the
owner of \texttt{foo}.

@@

\subsection{{\large\bf\textit{lastChange}} on Variable} 
In JavaScript, every frame has a scope chain and every available
variable in the frame comes from one of the scopes in the frame's
scope chain. Once the developer asks for the last change of a variable with name 
\texttt{foo} at the reproduction point, the debugger first determines the
variable's scope as follows: it iterates over the scopes in the scope
chain and the first scope which has a variable with the same name is
the variable's scope. There are five different scope types: global, local,
closure, \texttt{with} and \texttt{catch}. We explain these
cases in two groups.

@@

\begin{figure}[htp]
%\centering 
%\begin{algorithm}
\line(1,0){240}
\begin{algorithmic}%[frame=single]
\FOR{$q$ in $lastChange$ queries} 
 \FOR {$p$ is defined at the $q$ result}
   \IF {$p$ is a $lastChange$ on object property} 
     \STATE the property {\bf owner id} must be stored at $q$ change events. 
   \ELSIF {$p$ is a $lastChange$ on variable} 
     \STATE the variable {\bf scope id} must be stored at $q$ change events.
	 \ENDIF 
 \ENDFOR 
\ENDFOR

\end{algorithmic}
\line(1,0){240}
%\end{algorithm}
\caption{\textit{lastChange} queries dependency analysis.}
\label{fig:dependency-analysis}
\end{figure}
%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented\footnote[1]{http://code.google.com/p/querypoint-debugging} it in an extension to the Firebug
JavaScript debugger\footnote[2]{http://getfirebug.com}. %\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser. %\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface \cite{JSD} and 
\textit{Querypoint} is developed over this interface. Our prototype implements the four primitive operations in JavaScript 
using techniques which are cumbersome and comparatively slow to execute. However the JavaScript prototype is
easy to explore, change, and share with others for feedback. A professionally useful debugger 
would implement these primitive operations within the JavaScript engine.

\subsection{{\normalsize\bf\texttt{objectId()}} Operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}
\footnote[1]{The property name we used in the prototype is longer and more specific to reduce the chance of name conflicts.} .
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. 
%\footnote[3]{This use of a regular object property  might change
%the program behaviour. For example it will appear as an extra property in 
%\texttt{for(property in object)} loops. 
%The next version of Firefox (https://developer.mozilla.org/en/JavaScript/Reference\\/Global\_Objects/Object/defineProperty) provides \texttt{defineProperty}, 
%allowing a property to be added which is invisible to enumeration.} Note that \texttt{\_objectId} is
%not set for all objects but only those objects that need an id.

\subsection{{\normalsize\bf\texttt{setPropertyChangeHook()}} Operation}
The Firefox JavaScript engine supports watching property changes in an
object. Every object has a function \texttt{watch(propName, callback)}
which receives two parameters, a property name and a function.  Whenever
the property with the given name changes, the \texttt{callback} function is
called. The hook set by this function remains enabled even if the
property is deleted and defined again.

@@

The only data we have is the object creation location including the
file url and the line number and the goal is to get a reference to
this object. Although in most regular cases we have only one
statement and one object creation, there are cases where more
objects are created in the line. There is no simple way to
recognize the interesting object among these new objects. So instead
of one object, we monitor all new objects created in the line.

An object might be created by one of these statements: object
literal(\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statement (\texttt{function()}). 
By parsing the source code we can recognize the statements that
create an object. The next step is getting a reference to the new object.

@@

The new object can also be set as the property of a parent object 
(or array) inside an object (or array) literal. This case is also treated 
similar to previous case. The only difference is that the full path of property 
from the root parent in the local scope must be considered as the assignee statement. For
example in Figure~\ref{fig:objectCreation}(b), if the creation line
is 20, we keep \texttt{parent.child}. In this case stepping
must be continued until the end of the object literal (line 22) for getting a reference
to the created new object.

In cases where the new object is passed as an 
argument to a function (Figure~\ref{fig:objectCreation}(c)), we use step-in instead of step-over. 
To get a reference to the new object
it is enough to evaluate the corresponding argument inside the function. The other cases
where the program does not keep a reference to the created object (e.g., when a function object is
just called after its creation), are not in our interest.

%(or \texttt{this} if the object is being constructed by \texttt{new} statement)

Although this approach is successful in many ordinary cases, we can imagine 
cases where a more comprehensive analysis needed for the correct behaviour.
Consider the case where the new object is assigned to an expression like \texttt{a[++i]}. 
% (where i is a variable). 
%This expression is kept in the list                      %(Figure~\ref{fig:objectCreation}(d))
%of assignee statements to be evaluated later for getting a reference to the new object. 
Obviously, evaluating this statement doesn't return
a reference to the new object. Our prototype implementation does not handle these kinds of unusual cases yet.

Throughout this section we implicitly assumed that the object will be created at 
the same location in the next execution. If the assumption is
not true, once the execution reaches the reproduction point, it
reveals that the object has been created in a different location. This time,
prototype re-executes the program considering both locations as possible 
object creation locations.

@@

(a) Case 1:
 20  x.y = new MyClass();

(b) Case 2:
 19  parent = {
 20    child : {
 21      x : 5
 22  }}

@@

\end{lstlisting}

%(d) Case 4:
% 20 a[++i] = new myConstructor();

%\end{verbatim}
\caption{Examples of different cases in getting references to created JavaScript objects.}
\label{fig:objectCreation}
\end{figure}

\subsection{{\normalsize\bf\texttt{scopeId()}} Operation}
The prototype sets a breakpoint at the beginning of all code blocks needing
a scope id. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook is hit,  meaning a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval("var \_scopeId = 10")} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{{\normalsize\bf\texttt{setVariableChangeHook()}} Operation}
Variables defined in local, closure, and catch scopes are only changed in
their scope; that includes the defining scope and scopes nested whithin them.
For example in Figure~\ref{fig:js-closure}, variable \texttt{foo} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line where the variable is defined. In Figure~\ref{fig:js-closure}, 
if the execution is paused at line 17 and the
last change of \texttt{foo} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different: \texttt{foo} in 
\texttt{childOne} is a local variable
but in \texttt{childTwo} it is a closure variable.

@@

10  function main(){
11    var foo;
12    foo = ...;
13    function parent(){
14      foo = ...;
15      function childOne(){
16        var foo;
17        foo = ...;
18      }  
19      function childTwo(){
20        foo = ...;			      
21      }
22    }  
23  }    
\end{lstlisting}
%\end{verbatim}
\caption{Sample JavaScript code demonstrating local and closure variables.}
\label{fig:js-closure}
\end{figure}

\subsection{Re-Execution, Reproduction Point and Data Collection}
\textit{Querypoint} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypoint}
is also able to automatically create them from the first execution.

@@

Since we don't need to compare values across executions to implement \textit{lastChange}, we can get more information if we do compare:
 different points of last change on different executions will signal that the execution is not deterministic. Note the converse is not true, many non-deterministic 
programs will give identical results for \textit{lastChange}.
Consider the example in Figure~\ref{fig:nondeterministic-values}. 
Assume that at the first execution, the developer sees that the \texttt{a} value is \texttt{true} and asks
for the \textit{lastChange} of \texttt{a} at the breakpoint. In the re-execution, \texttt{a} is \texttt{false},
and \texttt{b} is \texttt{true} at the reproduction point. It means that \texttt{lastChange}
result shows line 10, which is a correct answer for this execution but not for the previous one.
When user asks for the \textit{lastChange} on variable \texttt{a}, debugger stores \texttt{a} value and compares
it to the \texttt{a} value in the next executions. So if this value is different, debugger informs the developer
that the \textit{lastChange} query is made on a different value.
% In cases that the value is not a primitive but an object,
%the primitive properties of the object examined by the developer are stored and compared.

@@

10 a = b = false;
20 if (random()) {a = true;} else {b = true;}
30 if (a || b) {bug(); /* breakpoint */}
\end{lstlisting}
%\end{verbatim}
\caption{\textit{lastChange} on non-deterministic values.}
\label{fig:nondeterministic-values}
\end{figure}

\subsection{Combination of {\large\bf\textit{lastChange}} and Breakpoint Debugging}
\label{sec:pausing}
Using \textit{lastChange} a developer can work backwards on the flow of data, but sometimes bugs are more obvious when
we watch control flow forwards. Consider for example,

@@

In the case that the execution is deterministic, the \textit{last-\ Change}
event index can be used as an index for a conditional breakpoint \cite{Boothe, Maruyama}. 
Moreover, this works in a non-deterministic execution if non-determinism
has no effect on the event index.  The debugger can easily set this conditional breakpoint
and reexecute the program. If the stream of change events differs in this re-execution the
developer can be warned that the conditional breakpoint may not be the \textit{lastChange} point.
However, if the stream of events is the same, we do not know that the 
conditional breakpoint matches the last 
change, this fact can only be verified at the reproduction point. These theoretical concerns are not
likely to cause significant practical problems: if the value at the conditional breakpoint is not suspicious, 
then the developer will know to simply return to \textit{lastChange} or move on to another tactic to find the bug.

@@

We designed the example programs to require two steps of \textit{lastChange}; 
all four users took more steps (9, 13, 12, and more than 30) to find the defect location with breakpoints (Figure~\ref{fig:userstudy}).  
To find locations for breakpoints two users scrolled through the source, 
another searched the text, a third set a lot of breakpoints to get an understand of the control flow. 
Based on our own experience we expect these strategies represent the kinds of approaches 
developers have available. These operations are time consuming and tedious. In contrast, all four users 
found the defect location with just two \textit{lastChange} operations.
% Based on our measurements,
%this makes \textit{lastChange} about three times faster than breakpoints for the navigation from 
%the reproduction point to the defect
We recognize that these programs were designed to highlight lastChange and many kinds of debugging 
issues have been hidden by the design of our tests. Nevertheless our results show that, when a defect 
relates to incorrect values and a developer recognizes this, then the operational 
mechanics of lastChange lead to the defect much more quickly than breakpoints.

@@

\begin{figure}[htp]
\centering 
\includegraphics[width=.48\textwidth]{10-userstudy.pdf}
\caption{The number of steps and time (in seconds) every developer spent on each program
before locating the defect. The highlighted numbers show cases where the developer was
permitted to use \textit{lastChange}.}
\label{fig:userstudy}
\end{figure}

@@

%In our own experience, \textit{lastChange} dramatically reduces the work in
%setting and removing breakpoints, so work to cause re-execution  become noticeable.
%Therefore automatic mechanisms
%for re-execution will be corresponding more valuable for debugging
%with \textit{lastChange} (Our \textit{Querypoint} prototype  
%implements a simple form of automatic 
%re-execution for both breakpoint and \textit{lastChange}).
